### 2024-11-20 paper.tex
Update paper.tex

%% based on templae/IEEEtran/bare_conf.tex
\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{url}
\usepackage{fancyvrb}
\usepackage{courier}
\usepackage{helvet}
\usepackage{tikz}
\usepackage{xcolor}	
\usepackage{pdfpages}
\usepackage{amsmath}                   % AMS math typesetting
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{float}
\usepackage[parfill]{parskip}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%\usepackage[usenames,dvipsnames]{color}
\newcommand{\todo}[1] {\textcolor{red}{\textbf{TODO: #1}}}
\newcommand{\lhnote}[1] {(\textbf{LH says: \textcolor{red}{#1}})}
%\newcommand{\fix}[1]{\textcolor{red}{#1}}
\newcommand{\status}[1] {\textcolor{magenta}{\textbf{Status: #1}}}
\newcommand{\punchline}[1]{\textbf{Punch line: #1}}
\newcommand{\question}[1] {\textcolor{blue}{\textbf{#1}}}
%\newcommand{\punchline}[1]{}

\renewcommand{\punchline}[1]{}
\renewcommand{\todo}[1]{}
\renewcommand{\lhnote}[1]{}
\renewcommand{\status}[1]{}

% saves you writing {} after \msa. we can change \msa to \xmsa easily
\usepackage{xspace}
\newcommand{\msa}{$MSA$\xspace} 
\newcommand{\msat}{$MSA^T$\xspace}
\newcommand{\cnc}{Component \& Connector\xspace}
\usepackage{graphics}
%\input{tags/listings.tex}

\usepackage{listings}
\lstset{
    numbers=left,
    numberstyle=\tiny,
    numbersep=5pt,
    xleftmargin=10pt,
    xrightmargin=8pt,
    frame=single,
    aboveskip=5pt,
    belowskip=-4pt,
    sensitive=true,
%    float=!t,
    breaklines=true,
    captionpos=b,
    tabsize=2,
    showstringspaces=false,
    basicstyle=\small\ttfamily,
    morecomment=[l]{//},
    morecomment=[s][\itshape]{/**}{*/}
}
\lstdefinelanguage{MCG}[]{Java}{
  morekeywords={grammar, options, token}
}

%Equalize Reference length
\IEEEtriggeratref{15}

%\usepackage[margin=1in]{geometry}
%\textwidth=6.5in
%\textheight=9in
%\setlength{\topmargin}{-.50in}
%\setlength{\oddsidemargin}{0in}

\begin{document}

\title{Eliminating Input-Based Attacks by Deriving Automated Encoders and Decoders from Context-free Grammars}

\author{\IEEEauthorblockN{Tobias Bieschke\textsuperscript{1},
Lars Hermerschmidt\textsuperscript{1,2},
Bernhard Rumpe\textsuperscript{1} and
Peter Stanchev\textsuperscript{1}}
\IEEEauthorblockA{Software Engineering, RWTH Aachen University, Germany\textsuperscript{2}\\
AXA Konzern AG, Germany\textsuperscript{2}
%Email: tobias.bieschke@rwth-aachen.de, hermerschmidt@se-rwth.de, rumpe@se-rwth.de, peter.stanchev@rwth-aachen.de}
}
}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

\maketitle

\IEEEpeerreviewmaketitle

\begin{abstract}
%context of this work
Software systems nowadays communicate via a number of complex languages.
This is often the cause of security vulnerabilities like arbitrary code execution, or injections.
Whereby injections such as cross-site scripting are widely known from textual languages such as HTML and JSON that constantly gain more popularity.
These systems use parsers to read input and unparsers write output, where these security vulnerabilities arise.
Therefore correct parsing and unparsing of messages is of the utmost importance when developing secure and reliable systems.
%problem with existing research
Part of the challenge developers face is to correctly encode data during unparsing and decode it during parsing.

%Therefoer this paper aims to
This paper presents McHammerCoder, an (un)parser and encoding generator supporting textual and binary languages.
%The aproache works like...
Those (un)parsers automatically apply the generated encoding, that is derived from the language's grammar.
%This results in
Therefore manually defining and applying encoding is not required to effectively prevent injections when using McHammerCoder.
%The benefit for the user is...
By specifying the communication language within a grammar, McHammerCoder provides developers with correct input and output handling code for their custom language.
\end{abstract}

\section{Introduction}
\punchline{String Concat is the problem}

%open two perspectives on the problem: application developer and language developer
\punchline{Why do developers use it}
it is easy to use because every programming language offers primitives with strings and bytes, so developers got used to it

\punchline{Related approaches}
why static code analysis failed to solve the problem: Developers would have to use another tool. They only use what they have to: the compiler

Dear compiler developers, let's have a talk
Dear everyone else, from now on we assume you know how to build a compiler.

\punchline{building SAST into compilers will not solve the problem}
\cite{rice}

\punchline{redesign the language}
remove string Concat

\punchline{the solution has to be as easy as string concatenation}

\punchline{summarize solution from both perspectives}

\section{Related Work}\label{sec:relwork}
\punchline{Lars bashing, how could you assume that developers will use your tool/library}

\section{Formal Problem Definition} \label{sec:problem}

All Software relies on input
devlo

It's the programming language job to prevent application developers from creating injection vulnerabilities.
Adding more sugger later will not make the cake any more tasty.

Problem to solve: Developers continue to develop software with injections

definition injection

language concepts of a programming language result in properties of running code
Example: Memory Safe Language

\punchline{define input safe language}
(string) concat free language -> code without infections

definition injection free: parse(unparse(t)) = t

for all outputs o=unparse(parse(o))

\punchline{definition shotgun unparser}
let k be a keyword of grammar G
w(k) = ...
unparse_concat(k) != unparse_lang(k)

$con=unparse_concat(k)$

case 1: let $con$ be syntactic incorrect
$parse(con)$ will fail and by that prevent injection

case 2: let $con$ be syntactic correct
$parse(con) != k$ will result in a different parse tree, which is an injection

Let's now tak a look at unparse_lang(k)
$lan = unparse_lang(k)$
$lan$ is always syntacticaly correct
$parse(lan) = k$ therefore there are no injections

unparse_concat ignores syntax -> may result in syntactic incorrect outputs, however correct 
unparse_lang uses grammar to create correct syntax

unparse_p1(i)=x
unparse_p2(parse_p2(x))=x

unparse_p1 != unparse_p2, since developers create these unparsers

\section{Noncat}\label{sec:mchc}

\section{Evaluation}
\section{Conclusion} \label{sec:conclusion}

\bibliographystyle{IEEEtran}
\bibliography{paper}
\end{document}