### 2024-11-22 paper.tex
Continue

%% based on templae/IEEEtran/bare_conf.tex
\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{url}
\usepackage{fancyvrb}
\usepackage{courier}
\usepackage{helvet}
\usepackage{tikz}
\usepackage{xcolor}	
\usepackage{pdfpages}
\usepackage{amsmath}                   % AMS math typesetting
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{float}
\usepackage[parfill]{parskip}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%\usepackage[usenames,dvipsnames]{color}
\newcommand{\todo}[1] {\textcolor{red}{\textbf{TODO: #1}}}
\newcommand{\lhnote}[1] {(\textbf{LH says: \textcolor{red}{#1}})}
%\newcommand{\fix}[1]{\textcolor{red}{#1}}
\newcommand{\status}[1] {\textcolor{magenta}{\textbf{Status: #1}}}
\newcommand{\punchline}[1]{\textbf{Punch line: #1}}
\newcommand{\question}[1] {\textcolor{blue}{\textbf{#1}}}
%\newcommand{\punchline}[1]{}

\renewcommand{\punchline}[1]{}
\renewcommand{\todo}[1]{}
\renewcommand{\lhnote}[1]{}
\renewcommand{\status}[1]{}

% saves you writing {} after \msa. we can change \msa to \xmsa easily
\usepackage{xspace}
\newcommand{\msa}{$MSA$\xspace} 
\newcommand{\msat}{$MSA^T$\xspace}
\newcommand{\cnc}{Component \& Connector\xspace}
\usepackage{graphics}
%\input{tags/listings.tex}

\usepackage{listings}
\lstset{
    numbers=left,
    numberstyle=\tiny,
    numbersep=5pt,
    xleftmargin=10pt,
    xrightmargin=8pt,
    frame=single,
    aboveskip=5pt,
    belowskip=-4pt,
    sensitive=true,
%    float=!t,
    breaklines=true,
    captionpos=b,
    tabsize=2,
    showstringspaces=false,
    basicstyle=\small\ttfamily,
    morecomment=[l]{//},
    morecomment=[s][\itshape]{/**}{*/}
}
\lstdefinelanguage{MCG}[]{Java}{
  morekeywords={grammar, options, token}
}

%Equalize Reference length
\IEEEtriggeratref{15}

%\usepackage[margin=1in]{geometry}
%\textwidth=6.5in
%\textheight=9in
%\setlength{\topmargin}{-.50in}
%\setlength{\oddsidemargin}{0in}

\begin{document}

\title{Input Safe Languages: Unparsing as first class citizen of programming languages}

\author{\IEEEauthorblockN{
Lars Hermerschmidt\textsuperscript{1} and
Mark Vollmary\textsuperscript{2}}
\IEEEauthorblockA{REWE digital, Germany\textsuperscript{1}\\
AXA Konzern AG, Germany\textsuperscript{2}}
}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

\maketitle

\IEEEpeerreviewmaketitle

\begin{abstract}
%context of this work
Software systems nowadays communicate via a number of complex languages.
This is often the cause of security vulnerabilities like arbitrary code execution, or injections.
Whereby injections such as cross-site scripting are widely known from textual languages such as HTML and JSON that constantly gain more popularity.
These systems use parsers to read input and unparsers write output, where these security vulnerabilities arise.
Therefore correct parsing and unparsing of messages is of the utmost importance when developing secure and reliable systems.
%problem with existing research
Part of the challenge developers face is to correctly encode data during unparsing and decode it during parsing.

%Therefoer this paper aims to
This paper presents McHammerCoder, an (un)parser and encoding generator supporting textual and binary languages.
%The aproache works like...
Those (un)parsers automatically apply the generated encoding, that is derived from the language's grammar.
%This results in
Therefore manually defining and applying encoding is not required to effectively prevent injections when using McHammerCoder.
%The benefit for the user is...
By specifying the communication language within a grammar, McHammerCoder provides developers with correct input and output handling code for their custom language.
\end{abstract}

\section{Introduction}
\punchline{String Concat is the problem}

%open two perspectives on the problem: application developer and language developer
\punchline{Why do developers use it}
it is easy to use because every programming language offers primitives with strings and bytes, so developers got used to it

\punchline{Related approaches}
why static code analysis failed to solve the problem: Developers would have to use another tool. They only use what they have to: the compiler

Dear compiler developers, let's have a talk
Dear everyone else, from now on we assume you know how to build a compiler.

\punchline{building SAST into compilers will not solve the problem}
\cite{rice}

\punchline{redesign the language}
remove string Concat

\punchline{the solution has to be as easy as string concatenation}

\punchline{summarize solution from both perspectives}

\section{Related Work}\label{sec:relwork}
\punchline{Lars bashing, how could you assume that developers will use your tool/library}
\todo{trusted types anschauen}

\section{Formal Problem Definition} \label{sec:problem}

All Software relies on input

It's the programming language job to prevent application developers from creating injection vulnerabilities.
Adding more sugar later will not make the cake any more tasty.

Problem to solve: Developers continue to develop software with injections

definition injection

language concepts of a programming language result in properties of running code
Example: Memory Safe Language

\punchline{memory safety is not the solution}

Memory safe languages only solve one part of the input problem. A memory safe parser can't protect from injections.
Parsers are still a problem but memory safe languages ​​have fixed many of the bad effects.

\punchline{define input safe language}
(string) concat free language -> code without infections

definition injection free: parse(unparse(t)) = t

for all outputs o=unparse(parse(o))

\punchline{definition shotgun unparser}
let k be a keyword of grammar G
w(k) = ...
unparse_concat(k) != unparse_lang(k)

$con=unparse_concat(k)$

case 1: let $con$ be syntactic incorrect
$parse(con)$ will fail and by that prevent injection

case 2: let $con$ be syntactic correct
$parse(con) != k$ will result in a different parse tree, which is an injection

Let's now tak a look at unparse_lang(k)
$lan = unparse_lang(k)$
$lan$ is always syntactically correct
$parse(lan) = k$ therefore there are no injections

unparse_concat ignores syntax -> may result in syntactic incorrect outputs, however correct 
unparse_lang uses grammar to create correct syntax

unparse_p1(i)=x
unparse_p2(parse_p2(x))=x

unparse_p1 != unparse_p2, since developers create these unparsers

\section{Noncat}\label{sec:mchc}

Ansatz: template engine
Wir haben das was echt schlechtes gebaut was Devs nicht annehmen.
Loesungsansatz fuer regulaere output Sprache.
Ergebnis: App Devs haben das Problem nciht da sie ein Framework nutzen was parser/unparser mitbringt ... aber bauchen csv in json ein :D
Learning: Devs merken nicht wenn sie eine Sprache selber definieren
Learning: Devs erwarten dass die Sprache oder das Framework das Problem fuer sie loest (aehnich wie bei memory safety)

\section{Evaluation}

\todo{Wie gross ist der Bias weil wir die template engine als library und nicht als eigene sprache implementiert haben?}
\todo{brauchen wir einen zweiten ansatz mit einer DSL?}
Wie evaluiert man bessere Loesungen?
Wie koennen Leute messen ob sie mit ihrer Loesung auf dem Richtigen weg sind?
Interviews machen und Fragen veroeffentilchen.
- Dev Stakeholder Akzeptanz
- Business Stakeholder (wie viel Geld wuerdest du ausgeben 80\% deiner vulnerabilities loszuwerden? Welchen Teil des Security Budget wird frei?)

\section{Conclusion} \label{sec:conclusion}

\bibliographystyle{IEEEtran}
\bibliography{paper}
\end{document}